---
title: Reading in shapefiles in R
author: Joe Brew and Raquel Catalão
fig_height: 2.6
fig_width: 4
output:
  html_document:
    toc: true
    toc_float: true
    theme: united
---

```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = FALSE}
# No scientific notation
options(scipen=999)

# Packages 
library(xtable)
library(knitr)
library(ggplot2) 
library(dplyr)

# Basic knitr options
opts_chunk$set(comment = NA, 
               echo = TRUE, 
               warning = FALSE, 
               message = FALSE, 
               error = TRUE, # Render report, even with errors
               cache = F)

```

```{r setup, include=FALSE, echo = FALSE}
library(knitr)
knit_engines$set(asis = function(options) {
  if (options$echo && options$eval) knit_child(text = options$code)
})
```

# Directories

Make a folder called "data" and put it on your desktop. Download and unzip the shapefiles you sent into that folder. Then open an Rscript and save it to your desktop also. Call it "code.R" (or something else - it doesn't matter).

Open your "code.R". Ensure that your "working directory" is your desktop. You can do this by running `getwd()`. If it is not your desktop, run `setwd("Path/To/Your/Desktop")` (and alter the preceding text).

# Packages

You'll need some packages.

```{r}
# Load the necessary libraries
library(rgdal)
library(raster)
library(sp)
library(leaflet)
```

If running the above gives you an error, it's because you need to install the `rgdal` package (or another one of the packages). Do so like this:

```{r, eval = FALSE}
install.packages('rgdal')
```

# Read in data

Now you can use the `readOGR` function to read in your data. You were sent both "clip" data and "moz" data - we'll read in both shapefiles.

```{r}
clip <- readOGR('data', 'FACILITIES_CLIP')
moz <- readOGR('data', 'FACILITIES_MOZ')
```

# Also get a Mozambican map

While we're at it, let's also get a Mozambican map to plot on top of:

```{r}
map <- getData('GADM', country = 'MOZ', level = 1)
```

# Standardize coordiantes

We need to standardize the coordinate and projection systems between our different data sources. We'll make everything latitude/longitude, for simplicity's sake.

```{r}
clip <- spTransform(clip, proj4string(map))
moz <- spTransform(moz, proj4string(map))
```


# Visualize the data


## Static visualization


Here are all the points in the "Moz" shapefile you sent.
```{r}
# Plot the mozambican map
plot(map)

# Add points for all of mozambique
points(moz, 
       col = adjustcolor('black', alpha.f = 0.6))
```

Here are all the points in the "clip" shapefile you sent.

```{r}
# Plot the mozambican map
plot(map)

# Add points for all of mozambique
points(clip, 
       col = adjustcolor('darkred', alpha.f = 0.6))
```

Here's a zoom-in of just the clip area.

```{r}
# Subset our map
sub_map <- map[map$NAME_1 %in% c('Maputo', 'Maputo City', 'Gaza'),]

# Plot the subset map
plot(sub_map)

# Add points for all of mozambique
points(clip, 
       col = adjustcolor('darkred', alpha.f = 0.6))
```

# Interactive map

Below is an interactive map of the clip area. Click on any point to see the location. Change the `addProviderTiles` line to make different kinds of maps (some examples are there, commented out).

```{r}
leaflet(sub_map) %>%
  # addProviderTiles('Stamen.Watercolor') %>% 
  # addProviderTiles("Esri.WorldImagery") %>%
  addProviderTiles('OpenStreetMap.BlackAndWhite') %>%
  addCircleMarkers(lng = clip$Longitude,
                  lat = clip$Latitude,
                  color = 'blue',
                  fillColor = 'blue',
                  radius = 5,
                  opacity = 0,
                  fillOpacity = 0.5,
                  popup = clip$Nome_US) 
```



Hope this works for you! Let me know if you run into any problems.

xx


